Incentives are the biggest barriers to widespread data sharing

### 2021-08-02 new-paper.Rmd
minor change of sentence to "In a study of applied computational research, less than 20% of software described in publications was able to be installed and run [@Collberg2014]"

---
title: Incentives are the biggest barriers to widespread data sharing
author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
csl: nature-communications.csl
header-includes: 
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

@@

Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence. Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by efforts like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015]. The bigger driver for this is more likely due to the growing popularity of data science as a discipline distinct from statistics [@Donoho2017]. This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools. Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data. Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular. Modern scientists now have access to many interactive computational frameworks (e.g. Jupyter, RMarkdown, Stencila, Observable) that can interface with kernels from dozens of programming languages. Beyond notebooks to run the code, a reader will also require access to the same software dependencies. In a study of applied computational research, [@Collberg2014] and colleagues were only able to install and run less than 20% of the software described in publications. In addition to locating software, one still has to locate the right versions. Dependency management is a controversial topic in programming communities with no easy solutions. While languages like JavaScript have solved this with package managers like npm, it is much harder to accomplish in others like Python and R which are currently supported by a patchwork of solutions. Although this problem is far from solved, technologies like Docker have alleviated some of these concerns. Researchers can share Dockerfiles to recreate computational environments [@Boettiger2015]. Services like Binder can then leverage Docker and make it possible to share the computational environments in which these notebooks were run. The barriers to verification are now low enough that trivial reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].

>>>>

---
title: Incentives are the biggest barriers to widespread data sharing
author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
citation_package: 'natbib'
csl: nature-communications.csl
header-includes: 
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

@@

Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence. Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by efforts like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015]. The bigger driver for this is more likely due to the growing popularity of data science as a discipline distinct from statistics [@Donoho2017]. This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools. Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data. Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular. Modern scientists now have access to many interactive computational frameworks (e.g. Jupyter, RMarkdown, Stencila, Observable) that can interface with kernels from dozens of programming languages. Beyond notebooks to run the code, a reader will also require access to the same software dependencies. In a study of applied computational research, less than 20% of software described in publications was able to be installed and run [@Collberg2014]. In addition to locating software, one still has to locate the right versions. Dependency management is a controversial topic in programming communities with no easy solutions. While languages like JavaScript have solved this with package managers like npm, it is much harder to accomplish in others like Python and R which are currently supported by a patchwork of solutions. Although this problem is far from solved, technologies like Docker have alleviated some of these concerns. Researchers can share Dockerfiles to recreate computational environments [@Boettiger2015]. Services like Binder can then leverage Docker and make it possible to share the computational environments in which these notebooks were run. The barriers to verification are now low enough that trivial reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].
